---
title: "Movie Tweetings Analysis"
author: "Vinh Phan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    number_section: true
    code_fold: hide
params:
  tbl_schema_url: '../sample_dataset/tbl_schema.txt'
  movies_url: '../sample_dataset/movies.dat'
  ratings_url: '../sample_dataset/ratings.dat'
  users_url: '../sample_dataset/users.dat'
runtime: shiny
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	out.width = "80%"
)
if(!require("pacman")){
  install.packages("pacman")
}
pacman::p_load(DT, readr, tidyr, dplyr, ggplot2, tmap, scales, leaflet, shiny, stringr, knitr, lubridate)

```



## Wrangling data

### Import datasets
```{r read_files, error=FALSE, warning=FALSE, include=FALSE}
# load data
data.tbl_schema <- read_csv(params$tbl_schema_url)
data.raw.users <- read_table(params$users_url, col_names = FALSE)
data.raw.ratings <- read_table(params$ratings_url, col_names = FALSE)
data.raw.movies <- read_table(params$movies_url, col_names = FALSE) # R doesnt allow read "::" due to limit 1 byte of separator

```


Let have a look at data structure of **Users.dat, Movies.dat, and Ratings.dat** respectively.
```{r check_data_structure, echo=TRUE}
# check data structure
glimpse(data.raw.users)
glimpse(data.raw.movies)
glimpse(data.raw.ratings)

```



### Format datasets
In this step, I transform raw data into suitable format which can be used for further analysis.

**Notes:** since "genre" is a nested data structure, so I separate it into rows
```{r wrangle_data, echo=TRUE}
# convert into dataframe
users_col <- data.tbl_schema$attributes[which(data.tbl_schema$dataset == 'users')]
users_col <- unlist(str_split(users_col, pattern = "::"))
movies_col <- data.tbl_schema$attributes[which(data.tbl_schema$dataset == 'movies')]
movies_col <- unlist(str_split(movies_col, pattern = "::"))
ratings_col <- data.tbl_schema$attributes[which(data.tbl_schema$dataset == 'ratings')]
ratings_col <- unlist(str_split(ratings_col, pattern = "::"))

# separate by "::"
data.users <- separate(data.raw.users, col = X1, into = users_col, sep = "::",remove = TRUE)
data.movies <- separate(data.raw.movies, col = X1, into = movies_col, sep = "#", remove = TRUE)
data.ratings <- separate(data.raw.ratings, col = X1, into = ratings_col, sep = "::", remove = TRUE)

# seprate each cell in genre to rows
data.movies <- separate_rows(data.movies, genre)

# transform timestamp to datetime format
data.ratings <- transform(data.ratings, rating = as.numeric(rating), rating_date = as.numeric(rating_timestamp))
data.ratings$rating_date <- as.POSIXct(data.ratings$rating_date, origin = lubridate::origin)
data.ratings <- data.ratings %>% select(user_id, movie_id, rating, rating_date)

```
Check Movies data after wrangling
```{r echo=TRUE}
# Movie data after cleaning
head(data.movies)

```

### Check missing and outliers
1. There are 66 movies missing genre. It's a minor number against 10.506 movies. So I decided to remove them. However, If we have more information about the data, and the system architect, we maybe able to trace them.
2. The datasets are clean, no outliers detected. As we can see in the boxplot below, the range of rating is between 0 and 10, and most of data points fall around 8.
```{r check_outliers, echo=TRUE, fig.cap = 'Figure 1.1: check outlier of rating feature in "ratings.dat" data.'}
# count empty genre
count(data.movies[data.movies$genre == "",])

# remove NA rows
data.movies <- data.movies[data.movies$genre != "",]

# check outlier of rating feature
boxplot(data.ratings$rating)

```


### Compute popular movies by genre, year by year based on user ratings
Zipf's power law states that high quantities have a low frequency and low quantities with high ranking have high frequency.

1. Popularity is estimated by a logarithmic function of Quanity and Quality of ratings
2. Quantity: is top 50 movies that have highest number of ratings
3. Quality: it top 50 movies that have highest ranting

```{r echo=TRUE}
# Set constant variables
TOP_K <- 50 # top k popular movies
HIGH_RATING <- 8 # any ratings less than 8 is not considered as High
GENRE <- sort(unique(data.movies$genre))

# join ratings vs movies
data.mov_rat <- data.ratings %>% left_join(data.movies, by = "movie_id") %>%
  mutate(rating_year = year(rating_date)) %>%
  select(user_id, movie_id, rating, rating_year, rating_date, movie_title, genre)

# Compute
comp.mov_rat <- data.mov_rat %>%
  group_by(movie_id, movie_title, genre, rating_year) %>%
  summarise(rat_number = n(), 
            rat_high_quality = sum(rating[rating >= HIGH_RATING]), 
            rat_mean_quality = round(mean(rating), 2),
            rat_log_num_high = round(sum(log(rat_number) + log(rat_high_quality)), 3))

# Select top_k movie / genre by quality (qua) or quantity (qty)
res.gen_mov.qty <- comp.mov_rat %>%
  group_by(genre) %>%
  top_n(TOP_K, rat_number) %>%
  arrange(genre, desc(rat_number)) %>%
  select(movie_id, movie_title, genre, rating_year, rat_number, rat_mean_quality, rat_log_num_high) %>%
  rename("Movie ID" = movie_id, "Title" = movie_title, "Genre" = genre, "Year Rated" = rating_year, 
         "# Ratings" = rat_number, "AVG Ratings" = rat_mean_quality, "Popularity" = rat_log_num_high)

res.gen_mov.qlt <- comp.mov_rat %>%
  group_by(genre) %>%
  top_n(TOP_K, rat_high_quality) %>%
  arrange(genre, desc(rat_high_quality)) %>%
  select(movie_id, movie_title, genre, rating_year, rat_number, rat_mean_quality, rat_log_num_high) %>%
  rename("Movie ID" = movie_id, "Title" = movie_title, "Genre" = genre, "Year Rated" = rating_year, 
         "# Ratings" = rat_number, "AVG Ratings" = rat_mean_quality, "Popularity" = rat_log_num_high)

res.gen_mov.popularity <- comp.mov_rat %>%
  group_by(genre) %>%
  top_n(TOP_K, rat_log_num_high) %>%
  arrange(genre, desc(rat_log_num_high)) %>%
  select(movie_id, movie_title, genre, rating_year, rat_number, rat_mean_quality, rat_log_num_high) %>%
  rename("Movie ID" = movie_id, "Title" = movie_title, "Genre" = genre, "Year Rated" = rating_year, 
         "# Ratings" = rat_number, "AVG Ratings" = rat_mean_quality, "Popularity" = rat_log_num_high)

```


## Visualization
I support 3 options for sorting, by Most Popular movies, Highest rating movies, and Highest view movies. Also, users can also filter by genre. Due to a limited data volume, I only have data of 2013, so the filter by year function is not support.

To identify which movies are more popular than others. I used logarithmic functions to normalize the data. I tried to balance the quantity and quality of ratings to give a best result possible.


```{r viz_render, echo=TRUE}
# Load dynamic data
tbl_data <- reactive({
  if(input$MyOption == 'Most Popular'){
    filter(res.gen_mov.popularity, Genre %in% input$MyGenre)
  }
  if(input$MyOption == 'Highest Rating'){
    filter(res.gen_mov.qlt, Genre %in% input$MyGenre)
  }
  else{
    filter(res.gen_mov.qty, Genre %in% input$MyGenre)
  }
})

# Render
fluidPage(
  titlePanel("Movie Popularity"),
  # Render dropdown box
  fluidRow(
    column(4, selectInput("MyOption", "Sort option: ", 
                                  choices = c("Most Popular" ,"Highest Rating", "Highest Views"),
                                  multiple = FALSE, selected = "Most Popular")
    ),
    column(4, selectInput("MyGenre", "Filter by Genre: ", 
                          choices = GENRE, multiple = FALSE, selected = "Action")
   )
  ),
  # Render data table
  renderDT({
    datatable(tbl_data(),
      options = list(pageLength = 10, autoWidth = TRUE)
    )
  })
)


```


## Conclusion

**As we can see in the table:**

- Top 5 most popular action movies are: *Man of Steel, Iron Man 3, Start Trek, Oblivion, and World War Z *. However, if I change sort option to Highest Rating, the result is different, *Mand of Steel, Iron Man 3, Star Trek, Fast & Furious 6, Pacific Rim*. 
  - We see a change in *Oblivision* by *Fast & Furious 6*. This is because *Oblivion* has higher number of ratings, but those ratings are lower than *Fast & Furious 6*.
  - In another word, Popularity is estimated not only by the number of ratings of that movie but also the quality of those ratings.


**Why this analysis is helpful**

- This analysis can give an initial movie recommendation for **Cold start visitors** who access the website in the first time. Since new visitors registers to the system and requests recommendations before providing any rating. Popular movies are very likely to be interesting to the majority of users and hence to serve them the initial stage of their interaction with a recommender system.






_________________
